---
title: "Lab 5"
author: "Tyler Edvall"
date: "2021-02-23"
output: github_document
---
Objectives: 

 -Rename variables

 -Create new variables

 -Summarize a dataset

 -Summarize a dataset by groups

 -Visualize the relationship between variables

  -Numerical and categorical variable
  -Two numerical variables
  -Three variables




1. Rename each variable so that it is all lower-case and uses an underscore _ instead of a period . in the name (the recommended coding style in the tidyverse style guide). Print the resulting table.

```{r}

# load packages -----------------------------

library(tidyverse)

# read data ---------------------------------

iris
iris <- as_tibble(iris)
iris

# question 1 -------------------------------

iris_new <- 
  rename(
    iris,
    sepal_length = Sepal.Length,
    sepal_width = Sepal.Width,
    petal_length = Petal.Length,
    petal_width = Petal.Width
  )

iris_new

```


2. Convert the four numerical variables from cm to mm by multiplying by 10. Print the resulting table.

```{r}
# question 2 -------------------------------

iris_numbers<- select(iris_new, sepal_length, sepal_width, petal_length, petal_width)
iris_numbers
mutate(iris_new, iris_numbers = iris_numbers * 10)

```


3. Calculate sepal area and petal area (area is equal to length multiplied by width). Print a table with only the variables sepal area, petal area, and species.

```{r}
# question 3 -------------------------------

Q3 <-mutate(iris_new, sepal_area = sepal_length * sepal_width, petal_area = petal_length * petal_width)
Q3
Q3.5 <- select(Q3, sepal_area, petal_area)
Q3.5

```

4. Calculate the following statistics for the entire dataset from the sepal length variable and print the resulting table:

  a. sample size
  b. maximum value
  c. minimum value
  d. range
  e. median
  f. first quartile (q1)
  g. third quartile (q2)
  h. inter-quartile range (iqr)

```{r}
# question 4 -------------------------------

summarize(
  iris_new,
  sampl_size = n(),
  min = min(sepal_length),
  max = max(sepal_length),
  range = max(sepal_length) - min(sepal_length),
  median = median(sepal_length),
  q1 = quantile(sepal_length, probs = 0.25),
  q2 = quantile(sepal_length, probs = 0.75),
  IQR = IQR(sepal_length)
)

```

5. Calculate the following statistics for each species from the petal width variable and print the resulting table:

  a. sample size
  b. mean
  c. standard deviation
  d. variance
  e. standard error of the mean
  f. approximate 95% confidence interval

```{r}
# question 5 -------------------------------

iris_grouped <-
  iris_new %>%
  group_by(Species) %>%
  summarize(
   sample_size = n(),
   mean = mean(petal_width),
   sd = sd(petal_width),
   var = var(petal_width),
   sem = mean / sqrt(sample_size),
    ci_upper = mean + 2 * sem,
    ci_lower = mean - 2 * sem
)
iris_grouped

```

6. Visualize the relationship between petal width and species using a strip plot.

```{r}
# question 6 -------------------------------

ggplot(data = iris_new) +
  geom_jitter(mapping = aes(x = Species, y = petal_width))

```

7. Starting with the previous graph, add the mean and 95% confidence interval for each species

```{r}
# question 7 -------------------------------

ggplot(data = iris_new) +
  geom_jitter(mapping = aes(x = Species, y = petal_width)) +
  geom_crossbar(
    data = iris_grouped, 
    mapping = aes(x = Species, y = mean, ymax = ci_upper, ymin = ci_lower),
    color = "red"
  )

```

8. Visualize the relationship between petal length, petal width, and species using a scatterplot. Map the two numerical variables to the x and y axes and map species to the color and shape aesthetics.

```{r}
# question 8 -------------------------------

ggplot(data = iris_new) +
  geom_point(mapping = aes(x = petal_length, y = petal_width, color = Species))

```

```{r session-info}
sessioninfo::session_info()
```


